---
output: github_document
---


```{r, include = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>",
#   fig.path = "man/figures/README-",
#   out.width = "100%"
# )
options(tibble.print_min = 5, tibble.print_max = 5)
```

# eas


## Overview

Package eas for [Einstein Aging Study](https://einsteinmed.org/departments/neurology/clinical-research-program/eas/)


The EAS is a longitudinal study of community-residing individuals, aged 70 and older, in the Bronx, New York, a racially and ethnically diverse urban setting. In May 2017, the EAS added the neuropsychological battery of the Uniform Data Set (UDSNB 3.0) to the in-person assessment battery. Below is the list of UDSNB 3.0 tests, the corresponding variables and item numbers on the UDSNB 3.0 form. [Click here for the defintion of each variable](https://files.alz.washington.edu/documentation/uds3-ivp-c2.pdf)


- Montreal Cognitive Assessment(MoCA)

- Craft Story 21 Recall(Immediate)

- Benson Complex Figure Copy

- Number Span Test: Forward

- Number Span Test: Background

- Category Fluency

- Trail Making Test 

- Craft Story 21 Recall(Delayed)

- Benson Complex Figure Recall

- Multilingual Naming Test (MINT) 

- Verbal Fluency: Phonemic Test


Current package provides a tool to calculate demographically adjusted z-scores and impairment indicators for UDS3 neuropsychological tests. You can specify if you want to use the norms from Einstein Aging Study(EAS) or National Alzheimer’s Coordinating Center(NACC).

- EAS norms

obtained from 225 cognitively normal older adults in EAS 

- NACC norms 

obtaied from comparable data of 5,031 participants in the NACC database 

Use norms regression estimates to generate gender, age, education, black race adjusted norms.
The mean is calculated as:

$$ \hat{Y}=\hat{b_0}+ \hat{b_1}*female + \hat{b_2}*(age-77) + \hat{b3}*(educyrs-16)+ \hat{b4}*black$$

For trail A1 and trail B1, as a higher score indicates a worse performance, z-score is calculated as 

$-(Y - \hat{Y})/\sigma$.

For all the other cognitive tests where a higher score indicates a better performance, z-score is calculated as 

$(Y - \hat{Y})/\sigma$.


Thus, an impairment can be defined as a z-score below a certain cutoff, for example, -1 or -1.5 are often used. 
Using a cutoff of -1 means that a subject is defined as impaired on this test if their performance is 1SD below the population mean. 


The indicator of impairment is generated as:
$$\text{impair} = 1 \;\;\;\text{if z-score} \le \text{cutoff}$$

## Installation 

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("JiyueQin/eas")
```


## Usage 






