---
output: github_document
---


```{r, include = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>",
#   fig.path = "man/figures/README-",
#   out.width = "100%"
# )
options(tibble.print_min = 5, tibble.print_max = 5, tibble.width = Inf)

```

# eas


## Overview

Package eas for [Einstein Aging Study](https://einsteinmed.org/departments/neurology/clinical-research-program/eas/)


The EAS is a longitudinal study of community-residing individuals, aged 70 and older, in the Bronx, New York, a racially and ethnically diverse urban setting. In May 2017, the EAS added the neuropsychological battery of the Uniform Data Set (UDSNB 3.0) to the in-person assessment battery. Below is the list of UDSNB 3.0 tests, the corresponding variables and item numbers on the UDSNB 3.0 form. [Click here for the UDSNB3.0 form.](https://files.alz.washington.edu/documentation/uds3-ivp-c2.pdf)


- Montreal Cognitive Assessment(MoCA)
   
  `mocascore`(1f)
  

- Craft Story 21 Recall(Immediate)
  
  `verbatimi`(3a), `paraphrasei`(3b)
  
- Benson Complex Figure Copy

  `bensonscorei`(4a)

- Number Span Test: Forward
  
  `numspancorf`(5a)

- Number Span Test: Background
  
  `numspancorb`(6a)
  

- Category Fluency

  `animals60sec`(7a), `vegetables60sec`(7b)

- Trail Making Test
  
  `tr_a1`(8a), `tr_b1`(8b)
  
- Craft Story 21 Recall(Delayed)

  `verbatimd`(9a), `paraphrased`(9b) 

- Benson Complex Figure Recall
 
  `bensonscored`(10a)

- Multilingual Naming Test (MINT) 

  `minttotal`(11a)

- Verbal Fluency: Phonemic Test
  
   `fwords60sec`(12a), `lwords60sec`(12d), `flword`(12g)

Current package provides a tool to calculate demographically adjusted z-scores and impairment indicators for UDS3 neuropsychological tests. You can specify if you want to use the norms from Einstein Aging Study(EAS) or National Alzheimer’s Coordinating Center(NACC).

- EAS norms

obtained from 225 cognitively normal older adults in EAS, use `?norms_coef_eas` for the documentation.  

- NACC norms 

obtaied from comparable data of 5,031 participants in the NACC database, use `?norms_coef_nacc` for the documentation. 



## Installation 

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("JiyueQin/eas")
```


## Usage 


```{r warning=F, message=F}
library(eas)
library(tidyverse)
# here is a sample hypothetical dataset used to calculate z-scores. You should prepare your dataset in this standard format.
head(sample_dat)
str(sample_dat)

# only include three test variables for easier display
sample_dat_small = sample_dat %>% select(female:black_race, minttotal, tr_a1, tr_b1)

# calculate z-scores and the impairment indicators for tr_a1 and tr_b1 with NACC norms and 1.5 SD to define impairment.
uds_z(sample_dat_small, c('tr_a1','tr_b1'), norms = 'nacc', impair_sd = 1.5)

# calculate the z-score and the impaiment indicator for minttotal with EAS norms and 1SD to define impairment
# Also output mean and sd estimates in addition to the z-scores and the impairment indicators.
uds_z(sample_dat_small, 'minttotal', norms = 'eas', impair_sd = 1, out_mean_sd  = T)

```



